Mendelian randomization for causal inference in Genome-wide association studies

### 2022-02-23 mr-review.Rmd
g-estimation added

\begin{eqnarray}
  \sum_i (y_i - f(y_i;\beta)) = 0\nonumber \\
  \sum_i z_i(y_i - f(y_i;\beta)) = 0
  (\#eq:gmm)
\end{eqnarray}

The GMM estimate is the value of the vector $\beta$ that sastisfies Equation \@ref(eq:gmm), where $f(\mathbb{E}(Y|\text{do}(X = x))$

@@

Specifically, @burgess_mendelian_2013 combined causal effect (ratio) estimates from $M$ independent SNPs by using Equation \@ref(eq:ivw).

>>>>

\begin{eqnarray}
  \sum_i (y_i - f(x_i;\beta)) = 0\nonumber \\
  \sum_i z_i(y_i - f(x_i;\beta)) = 0
  (\#eq:gmm)
\end{eqnarray}

The GMM estimate is the value of the vector $\beta$ that satisfies Equation \@ref(eq:gmm), where $f(x_i; \beta) = \mathbb{E}(Y|\text{do}(X = x))$. @palmer_instrumental_2011 developed numerical methods for obtaining estimates from Equations \@ref(eq:gmm).

Another semiparametric estimation method is G-estimation of a structural mean model [@burgess_review_2017;@robins_new_1986;@greenland_estimating_2008].  
We follow @burgess_review_2017 by defining the potential outcome $Y(x)$ as the outcome value that we would have observed had we set the exposure value $X$ to $x$. For example, $Y(0)$ denotes the observed outcome had we set the exposure to zero instead of its observed value $x$. The structural mean model for a continuous outcome is displayed in Equation \@ref(eq:smm).

\begin{equation}
\mathbb{E}(Y(x) - Y(0) | X = x, Z = z) = \beta_1 x
  (\#eq:smm)
\end{equation}

The causal effect parameter is $\beta_1$. @burgess_review_2017 derives the estimating equations for $\beta_1$, after noting that the conditional expectation, $\mathbb{E}(Y(0)| X = x, Z = z)$ is independent of $Z$ and reasoning that the causal effect is that value of $\beta_1$ that yields zero covariance between $Z$ and $\mathbb{E}(Y(0)| X = x, Z = z)$ (Equations \@ref(eq:smm1)).

\begin{equation}
\sum_i (z_{ik} - \bar z_{\cdot k})(y_i - \beta_1 x_i) = 0
  (\#eq:smm1)
\end{equation}

Note that $k$ indexes SNP instruments and ranges from 1 to $K$.

Specifically, @burgess_mendelian_2013 combined causal effect ratio estimates from $M$ independent SNPs by using Equation \@ref(eq:ivw).

@@
